The Effects of the NBA COVID Bubble on the NBA Playoffs

### 2020-11-08 Manuscript/NBABubbleAnalysis.Rmd
Update NBABubbleAnalysis.Rmd

# Data {#sec:data}
The 2020 bubble provides a new and exciting opportunity to study home-court
advantage for the NBA. Unlike college basketball, aside from a few 
exhibition/preseason games, the NBA always has a home and away team. So, for
the first time in NBA history the bubble allows NBA home and away performance
to be compared vs a control/neutral field. The NBA bubble, as a neutral court,
removed all 4 possible game location factors impacting home-court advantage 
hypothesized by @Carron1992. The NBA bubble featured 8 seeding games then a
standard playoff format. The focus of this study was the play during the playoff
games since it followed the standard playoff format and can easily be compared
back to other playoffs. For this study, the 2020 playoffs were compared against
the three previous playoffs collectively. It is important to make sure that any
observed differences are truly from a home-court advantage. So, in order to control
for the changing play style of the NBA, we will limit the study to 2020 vs 2017-2019.
The reason being the faster pace play and more common use of the three-point shot 
in modern basketball. If we used data from say 10 years ago, or earlier, observed
differences may not be from effects of the NBA bubble, but rather from the effects
of drastic changes in the style of play between the seasons. However, basketball
evolves slow enough that we can reasonably assume 2017-2019 are at least very 
close in pace and playing style to 2020.

Data was collected from the official NBA website. The main variables of interest
are whether or not the home team won, scoring totals for home and away teams, and
shooting percentages for home and away teams on two pointers, three pointers, and
free throws. These variables were very popular and frequently used in the related
literature discussed earlier. While many other measures could be used for measuring
the outcome of the game and team performance, scoring seemed to be the most 
important. The winner of a game is determined by who scores more points. So, 
naturally when studying the outcome of winning, changes in scoring and scoring 
efficiency are important to understand.

@@

# Methods {#sec:methods}

>>>>

# Data {#sec:data}
Data was collected from the official NBA website. The main variables of interest
are whether or not the home team won, scoring totals for home and away teams, and
shooting percentages for home and away teams on two pointers, three pointers, and
free throws. These variables were very popular and frequently used in the related
literature discussed earlier. While many other measures could be used for measuring
the outcome of the game and team performance, scoring seemed to be the most 
important. The winner of a game is determined by who scores more points. So, 
naturally when studying the outcome of winning, changes in scoring and scoring 
efficiency are important to understand.

@@

# Methods {#sec:methods}
The 2020 bubble provides a new and exciting opportunity to study home-court
advantage for the NBA. Unlike college basketball, aside from a few 
exhibition/preseason games, the NBA always has a home and away team. So, for
the first time in NBA history the bubble allows NBA home and away performance
to be compared vs a control/neutral field. The NBA bubble, as a neutral court,
removed all 4 possible game location factors impacting home-court advantage 
hypothesized by @Carron1992. The NBA bubble featured 8 seeding games then a
standard playoff format. The focus of this study was the play during the playoff
games since it followed the standard playoff format and can easily be compared
back to other playoffs. For this study, the 2020 playoffs were compared against
the three previous playoffs collectively. It is important to make sure that any
observed differences are truly from a home-court advantage. So, in order to control
for the changing play style of the NBA, we will limit the study to 2020 vs 2017-2019.
The reason being the faster pace play and more common use of the three-point shot 
in modern basketball. If we used data from say 10 years ago, or earlier, observed
differences may not be from effects of the NBA bubble, but rather from the effects
of drastic changes in the style of play between the seasons. However, basketball
evolves slow enough that we can reasonably assume 2017-2019 are at least very 
close in pace and playing style to 2020.

Comparisons between 2020 and 2017-19 home and away teams were made on home team
winning percentage, total team scoring and two-point, three-point and free throw
shooting. Comparing the differences in these metrics for home and away teams in
2020 vs previous years will provide valuable insights to the understanding of 
home-court advantage. We can see how going on the road may negatively impact 
away performance and how playing at home may positively impact home performance.
If there are differences in scoring for home or away teams the differences can
be used to show how home-court advantage affects overall performance of home
and away teams. While testing for differences in shooting will provide added
context for how home-court advantage specifically affects performance. 
Shooting percentages aren’t the only possible metrics affected by home-court
advantage, but they are the most obvious and certainly an important one. In
order to compare these differences simple two-sample z-tests were conducted.
Test for differences in means was conducted for comparing scoring totals. 
Tests for differences in proportions were conducted for comparing winning
percentage and the three different shooting percentages.
